Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-03-15 overview.tex
partial commit

%GB: The division of this subsection into subsubsections is somewhat arbitrary.
%GB: I wanted to do it at first in the same divisions as section 3, but this is not really
%GB: possible because there is no "AMR" equation.  Maybe somebody can think 
%GB: of a clever way to fix this.

@@

\subsubsection{Hydrodynamics and Gravity}

@@

\begin{eqnarray} 
\frac{\partial \rho}{\partial t} 
          + \frac{1}{a} \vecv \cdot \nabla \rho
    & = & - \frac{1}{a} \rho \nabla \cdot \vecv, 
        \label{eq:mass} \\
%
\frac{\partial \vecv}{\partial t}  
          + \frac{1}{a} ( \vecv \cdot \nabla ) \vecv
    & = & - \frac{\dot{a}}{a} \vecv 
          - \frac{1}{a \rho} \nabla p
          - \frac{1}{a} \nabla \phi, 
        \label{eq:momentum} \\
%
\frac{\partial \Ep}{\partial t} 
         + \frac{1}{a} \vecv \cdot \nabla E
   & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rho} + {\vecv \cdot \vecv} )
         - \frac{1}{a \rho } \nabla \cdot (p \vecv) 
%\nonumber   & &   %uncomment for 2-column
         - \frac{1}{a} \vecv \cdot \nabla \phi 
                + \Gamma - \Lambda.
       \label{eq:total_energy} 
\end{eqnarray}

%\begin{eqnarray} 
%\frac{\partial \rho}{\partial t} 
%          + \vecv \cdot \nabla \rho
%    & = & - \rho \nabla \cdot \vecv,
%        \label{eq:mass} \\
%%
%\frac{\partial \vecv }{ \partial t} 
%          + ( \vecv \cdot \nabla ) \vecv 
%    & = & - \frac{1}{\rho} \nabla p
%          - \nabla \phi,
%        \label{eq:momentum} \\
%%
% \frac{\partial E}{\partial t} + \vecv \cdot \nabla E
%    & = & - \frac{1}{\rho} \nabla \cdot (p \vecv) 
%          - \vecv \cdot \nabla \phi
%%\nonumber          \\  & &     % uncomment for 2-column
%     + \Gamma - \Lambda
%        \label{eq:total_energy}
%\end{eqnarray}
%
Here $E$, $e$, $\rho$, $\vecv$ and $p$ are the comoving total specific fluid energy, specific thermal energy, baryonic density, velocity and pressure, respectively.  
%Here we use primes to indicate variables defined in the absolute frame of reference (to distinguish them from a comoving frame).  
The first equation represents conservation of mass, the second, conservation of momentum, and the third, conservation of total fluid (kinetic plus thermal) energy.  They are respectively, the first, second and third moments of the Boltzmann equation. 
Terms representing radiative cooling ($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the energy equation (\ref{eq:total_energy}).

The total fluid specific energy ($E$) is given by:
\begin{equation}
E =  e + \frac{1}{2} \vecv \cdot \vecv.
        \label{eq:total_energy_def}
\end{equation}
It is useful to solve the equation of total energy because doing so enforces local energy conservation.

@@

\subsubsection{Godonuv MUSCL with Dednet (HD and MHD)}

@@

The current implementation of self-gravity in \enzo uses a Fast Fourier Technique \citep{Hockney88} to solve Poisson's equation on the top (level 0) grid.  The advantage of this is that it naturally allows both periodic and isolated boundary conditions for the gravity, choices which are very common in astrophysics and cosmology.  On subgrids, we interpolate the boundary conditions from that grid's parent (either the level 0 grid or some other subgrid).  Poisson's equation is then solved using a multigrid technique.  In \enzo\, self-gravity is optional.  There are also a number of options for static gravitational fields.

>>>>

@@

\subsubsection{Magnetohydrodynamics and Gravity}

@@

\begin{eqnarray} 
\frac{\partial \rho}{\partial t} 
          + \frac{1}{a} \nabla \cdot (\rho \vecv) & = & 0
        \label{eq:mass} \\
%
\frac{\partial \rho \vecv}{\partial t}  
   + \frac{1}{a} \nabla \cdot \left(\rho \vecv \vecv + \myvec{I} p - \vecB \vecB \right) & = &
    - \frac{\dot{a}}{a} \rho \vecv - \frac{1}{a} \rho \nabla \phi
    %
%          + \frac{1}{a} ( \vecv \cdot \nabla ) \vecv
%    & = & - \frac{\dot{a}}{a} \vecv 
%          - \frac{1}{a \rho} \nabla p
%          - \frac{1}{a \rho} (\nalba \times \vecB) \times \vecB
%          - \frac{1}{a} \nabla \phi, 
        \label{eq:momentum} \\
%
\frac{\partial E \rho}{\partial t} 
+ \frac{1}{a} \nabla \cdot \left[ (E + p) \vecv - (\vecB \cdot \vecv) \vecB \right] & = &
     - \frac{2\dot{a}}{a} E - \frac{1}{a} \vecv \cdot \nabla \phi + \frac{\dot{a}}{a} \frac{\vecB \cdot \vecB}{2}
%
%         + \frac{1}{a} \vecv \cdot \nabla E
 %  & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rho} + {\vecv \cdot \vecv} )
  %       - \frac{1}{a \rho } \nabla \cdot (p \vecv) 
%%\nonumber   & &   %uncomment for 2-column
 %        - \frac{1}{a} \vecv \cdot \nabla \phi 
 %               + \Gamma - \Lambda.
       \label{eq:total_energy}  \\
%
\frac{\partial \vecB}{\partial t} & = & \frac{1}{a}  \nabla \times (\vecv \times \vecB) - \frac{\dot{a}}{2a} \vecB \label{eq:induction}
\end{eqnarray}

%\begin{eqnarray} 
%\frac{\partial \rho}{\partial t} 
%          + \vecv \cdot \nabla \rho
%    & = & - \rho \nabla \cdot \vecv,
%        \label{eq:mass} \\
%%
%\frac{\partial \vecv }{ \partial t} 
%          + ( \vecv \cdot \nabla ) \vecv 
%    & = & - \frac{1}{\rho} \nabla p
%          - \nabla \phi,
%        \label{eq:momentum} \\
%%
% \frac{\partial E}{\partial t} + \vecv \cdot \nabla E
%    & = & - \frac{1}{\rho} \nabla \cdot (p \vecv) 
%          - \vecv \cdot \nabla \phi
%%\nonumber          \\  & &     % uncomment for 2-column
%     + \Gamma - \Lambda
%        \label{eq:total_energy}
%\end{eqnarray}
%
Here $E$, $e$, $\rho$, $\vecv$, $\vecB$, and $p$ are the comoving total fluid energy, specific thermal energy, baryonic density, velocity, magnetic field strength and pressure, respectively.  $I$ is the identify matrix and we use a set of units for the magnetic field such that magnetic xxx is unity.
%Here we use primes to indicate variables defined in the absolute frame of reference (to distinguish them from a comoving frame).  
The first equation represents conservation of mass, the second, conservation of momentum, and the third, conservation of total fluid (kinetic plus thermal) energy.  They are respectively, the first, second and third moments of the Boltzmann equation. 
Terms representing radiative cooling ($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the energy equation (\ref{eq:total_energy}).

The total fluid specific energy ($E$) is given by:
\begin{equation}
E =  e + \frac{1}{2} \vecv \cdot \vecv + \vecB \cdot \vecB
        \label{eq:total_energy_def}
\end{equation}
It is useful to solve the equation of total energy because doing so enforces local energy conservation.

@@

\subsubsection{Godonov MUSCL with Dedner (HD and MHD)}

@@

The current implementation of self-gravity in \enzo uses a Fast Fourier Technique \citep{Hockney88} to solve Poisson's equation on the top (level 0) grid.  The advantage of this is that it naturally allows both periodic and isolated boundary conditions for the gravity, choices which are very common in astrophysics and cosmology.  On subgrids, we interpolate the boundary conditions from that grid's parent (either the level 0 grid or some other subgrid).  The Poisson equation is then solved using a multigrid technique.  In \enzo\, self-gravity is optional.  There are also a number of options for static gravitational fields.
